---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(pmxhelpr)
```

```{r plot-data, eval=T, echo=F, include=F}
plot_data <- data_sad %>%
  filter(EVID == 0, CMT %in% c(1,2), FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

`plot_dvtime()` is a convenience wrapper around `ggplot` for  visualizion of a 
continuous dependent variable (*DV*) over time (*TIME*). The function accepts 
NONMEM-format datasets directly with internal processing (e.g., BLQ handling, 
summary statistics, dose-normalization) to ensure consistency in exploratory 
analyses of PK/PD data.

The function is optimized for datasets containing a nominal time variable
(*NTIME*), representing protocol-specified times of sample collection. This 
variable is used to bin data for calculation of summary statistics.

```{r plot-dvtime, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
plot_dvtime(data = plot_data, dv_var = DV, time_var = TIME, ntime_var = NTIME,
            col_var = Dose, cent = "mean")
```

{{< pagebreak >}}

A caption prints specifying the central tendency measure calculated and 
visualized on the plot, which can be silenced with `show_caption = FALSE`.

Arguments specifying dataset columns can be passed to the function as 
bare names or strings. The returned plot is a `ggplot` object, 
which can be further modified by adding additional `ggplot` layers.

```{r plot-dvtime-lin, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
plot_dvtime(data = plot_data, col_var = "Dose") +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)")
```

{{< pagebreak >}}

`plot_dvtime()` uses `ggplot2::stat_summary()` to calculate summary statistics.
This function performs calculations *after* axis transformation layers, such 
that`"mean"` returns the geometric mean for log-scale y-axes specified 
with `ggplot2::scale_y_log()`. 

This nuance is often overlooked by pharmacometricians when describing their own
figures, since the argument they supplied to the function was `"mean"`.

This silent behavior is made explicit in `plot_dvtime()` via the `log_y` 
argument, which prints a caption updates to indicate to the user that the 
summary statistic has changed along with the axis. 

```{r plot-dvtime-log, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
plot_dvtime(data = plot_data, col_var = Dose, log_y = TRUE) +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)")
```
