---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
options(scipen = 999)
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(patchwork)
library(mrgsolve)
library(pmxhelpr)
```

```{r data-sad-pk, eval=T, echo=F, include=F}
data_sad_pk <- data_sad %>% 
  filter(CMT %in% c(1,2)) %>% 
  filter(FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

```{r model-pk, eval=T, echo=F, include=F}
model <- model_mread_load("pkmodel")
```

```{r df_mrgsim_replicate, eval=T, echo=F, include=F}
simout <- df_mrgsim_replicate(data = data_sad_pk,
                     model = model,
                     replicates = 500,
                     carry_out = c("DOSE", "FOOD", "BLQ", "LLOQ"),
                     recover  = c("PART", "Dose"))
```

The VPC plotting toolchain for continuous longitudinal data within `pmxhelpr`
consists of a summary function, `df_vpcstats()`, and two dual-mode plotting 
functions, `plot_vpc_cont()` and `plot_vpc_cens()`. The summary `vpc_stats`
object contains all relevant summary statistics within each bin. Plotting 
functions can perform one-step summary statistic calculation and plotting, or
they can use `vpc_stats` objects in a two-step workflow to generate multiple 
plot types with identical grouping.

The seminal paper introducing the pcVPC method (PMID: 21302010) stated the 
following  regarding the handing of BLQ data in these diagnostics:

"Observations censored due to limit of detection will therefore need to be 
handled in the same way as previously suggested for drop-out censoring of 
observations (22). This approach is based on omitting censored observations 
(e.g., BQL) in a similar manner for both the observed and simulated datasets 
before calculating the VPC statistics. In this case, the interpretation of the
pcVPC/pvcVPC for the continuous observations will be heavily dependent on an 
accurate prediction of the data censoring. Therefore, diagnostics ensuring 
accurate prediction of censoring frequency (e.g., percentage of observation 
below LLOQ) versus the independent variable will be important before assessing 
the pcVPC/pvcVPC for the continuous observations."

`pmxhelpr` is optimized for BLQ handling in VPCs and pcVPCs, with functionality
to support the pcVPC workflow described in the paper above: 
  
1. Assessment of predictions of the censored data range with `plot_vpc_cens()`
2. Assessment of predictions of the continuous data range with `plot_vpc_cont()`
     
We will reuse the `vpc_stats` summary object to generate plots of both the 
censored and continuous data ranges saving on computation.

```{r df-vpcstats-dose, warning=FALSE}
vpcstats_ntime_dose <- df_vpcstats(data = simout, strat_var = Dose)
vpcstats_ntime <- df_vpcstats(data = simout)
```

`plot_vpc_cens()` is the companion diagnostic to `plot_vpc_cont()`, which
evaluates the predictive ability of the model for the censored data range.
It plots `obs_prop_blq` (the per-bin observed BLQ proportion) overlaid on the 
non-parametric confidence interval of `sim_prop_blq` across replicates (90% CI
by default) from the `vpc_stats` summary object. 

Let's use our precomputed `vpc_stats` object to generate BOTH censored and
continuous range VPCs, confirming adequate predictive ability by dose for both
data types before generating a pcVPC pooled across doses. 

{{< pagebreak >}}

For this example, we will suppress the replicate count in the continuous range
plot with `show_rep=FALSE` and retain it on the censored range plot and use 
`patchwork` to combine the two plots into a single image.

```{r plot-vpc-cens-dose, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_cont <- plot_vpc_cont(vpcstats_ntime_dose, show_rep = FALSE) + 
  scale_x_continuous(breaks = seq(0, 168, 24), limits = c(0, 168)) + 
  scale_y_log10(guide = "axis_logticks") + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") 

plot_cens <- plot_vpc_cens(vpcstats_ntime_dose) + 
  scale_x_continuous(breaks = seq(0, 168, 24),limits = c(0, 168)) + 
  labs(y = "Proportion BLQ", x = "Time (hours)") 
```


```{r plot-vpc-cont-dose-log, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
plot_cont / plot_cens
```

{{< pagebreak >}}

Now that we have verified that the model is adequately predictive of the 
censored data range across doses using `plot_vpc_cens()`, we can generate a 
single pcVPC plot for the continuous data range pooled across doses. 

For good measure, we will repeat our patchwork procedure here to show BOTH the
censored and continuous data pooled diagnostics together in one plot! 

NOTE: The predictive ability for the censored data range should *ALWAYS* be 
verified by dose, as the pattern of BLQ data is dependent on dose intensity.

```{r plot-pcvpc-cont-log, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_all_cont <- plot_vpc_cont(vpcstats_ntime, pcvpc = TRUE, show_rep=FALSE) + 
  scale_y_log10(guide = "axis_logticks") + 
  scale_x_continuous(breaks = seq(0, 168, 24), limits = c(0, 168)) + 
  labs(y = "PRED-corr. Conc. (ng/mL)", x = "Time (hours)") 

plot_all_cens <- plot_vpc_cens(vpcstats_ntime) + 
  scale_x_continuous(breaks = seq(0, 168, 24), limits = c(0, 168)) + 
  labs(y = "Prop. BLQ", x = "Time (hours)") 

plot_all_cont / plot_all_cens
```
