---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(patchwork)
library(pmxhelpr)
```

```{r plot-data, eval=T, echo=F, include=F}
plot_data <- data_sad %>%
  filter(EVID == 0, CMT %in% c(1,2), FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

`plot_dvtime()` is a convenience wrapper around `ggplot` for  visualization of a 
continuous dependent variable (*DV*) over time (*TIME*). The function accepts 
NONMEM-format datasets directly with internal processing (e.g., BLQ handling, 
summary statistics, dose-normalization) to ensure consistency in exploratory 
analyses of PK/PD data.

The function is optimized for datasets containing a nominal time variable
(*NTIME*), representing protocol-specified times of sample collection. This 
variable is used to bin data for calculation of summary statistics specified
by the `cent` argument. Options include:

  + `"none"`: No central tendency layer
  + `"mean"`: Mean after axis transformation (default).
  + `"mean_sdl"`: Mean +/- one standard deviation after axis transformation
  + `"mean_sdl_upper"`: Mean + one standard deviation after axis transformation
  + `"median"`: Median (50th percentile)
  + `"median_iqr"`: Median +/- one quartile (inter-quartile range: 
                      25th to 75h percentiles)
  + `"median_iqr_upper"`: Median + one quartile
  
One may want to visualize only the mean +/- standard devation without the
added visual noise of the data points. This can be accomplished by removing
the points using *theme*. 

```{r plot-dvtime-meansd-log, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
plot_dvtime(data = plot_data, col_var = "Dose", cent = "mean_sdl",
            log_y = TRUE,
            theme = plot_dvtime_theme(obs_point = pmx_point(alpha = 0))) +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)")
```

{{< pagebreak >}}

On the linear scale, the lower standard deviation error bar may cross the 
zero threshold, depending on the variability in the underlying data. 

The `"mean_sdl_upper"` option exists to handle this scenario. Because the value 
of concentration cannot cross zero, we can suppress the lower error bars and 
visualize only the upper error bar to reduce visuale noise.

Let's take a look at the outputs of `cent = "mean_sdl"` and 
`cent = "mean_sdl_upper"` side-by-side. We will suppress the captions in favor
of titles, and gather our identical legends together in the 
bottom of the plot using `patchwork::plot_layout()` and `ggplot::theme()`. 

```{r plot-dvtime-mean-sd-lin,  message=FALSE, warning=FALSE}
plot_sdl <- 
  plot_dvtime(data = plot_data, col_var = "Dose", cent = "mean_sdl", 
              show_caption = FALSE,
              theme = plot_dvtime_theme(obs_point = pmx_point(alpha = 0))) +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)", title = "Mean +/- SD")
```

```{r plot-dvtime-mean-sdup-lin, message=FALSE, warning=FALSE}
plot_sdl_upper <- 
  plot_dvtime(data = plot_data, col_var = "Dose", cent = "mean_sdl_upper",
              show_caption = FALSE,
              theme = plot_dvtime_theme(obs_point = pmx_point(alpha = 0))) +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)", title = "Mean + SD")
```

```{r plot-dvtime-mean-sdl-patch, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
plot_sdl + plot_sdl_upper + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

{{< pagebreak >}}

Perhaps we want to visualize central tendency and individual trends without
making any distributional assumptions. The non-parametric central tendency
measure requested with `cent = "median"`. 

We can specify an additional argument `id_var`, which denotes the
identifier variable in `data`. This argument may  be passed as a bare name
or a string. When `id_var` is specified, it is passed to the  *group* aesthetic 
in `ggplot` and used to connect longitudinal repeated measures within an 
individual (e.g., spaghetti plots). 

```{r plot-dvtime-med-id, message=FALSE, warning=FALSE}
plot_med_id <- 
  plot_dvtime(data = plot_data, col_var = "Dose", cent = "median", id_var = ID,
            log_y = TRUE, show_caption = FALSE,
            theme = plot_dvtime_theme(obs_point = pmx_point(alpha = 0))) +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)", title = "Median (Individual)")
```

```{r plot-dvtime-id,  message=FALSE, warning=FALSE}
plot_id <- 
  plot_dvtime(data = plot_data, col_var = "Dose", cent = "none", id_var = ID,
            log_y = TRUE, show_caption = FALSE,
            theme = plot_dvtime_theme(obs_point = pmx_point(alpha = 0))) +
  scale_x_continuous(breaks = seq(0, 168, 24)) +
  labs(y = "Concentration (ng/mL)", x = "Time (hours)", title = "Individual") + 
  theme(legend.position = "none")
```

```{r plot-dvtime-med-id-patch,  message=FALSE, warning=FALSE}
plot_id + plot_med_id + plot_layout(guides = "collect")
```