---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(patchwork)
library(pmxhelpr)
```

```{r plot-data, eval=T, echo=F, include=F}
plot_data <- data_sad %>%
  filter(EVID == 0, CMT %in% c(1,2), FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

`plot_dvtime()` is a convenience wrapper around `ggplot` for  visualization of a 
continuous dependent variable (*DV*) over time (*TIME*). The function accepts 
NONMEM-format datasets directly with internal processing (e.g., BLQ handling, 
summary statistics, dose-normalization) to ensure consistency in exploratory 
analyses of PK/PD data.

Imputation of data censored below the assay lower limit of quantification 
(BLQ) is handled in a standardized manner, without modifying the underlying 
data for visualization with the arguments `loq` and `loq_method`.

The `loq` argument specifies the numeric value of the LLOQ. This value can also
be inherited row-wise *if* the variable `LLOQ` is present in `data`.

The `loq_method` argument specifies the imputation method for missing 
observations (EVID = 0, MDV = 1).

  + `0` or `"none"`: No handling. (default)
  + `1` or `"postdose"`: 0 when `TIME` <= 0 and 1/2 x `loq` when `TIME` > 0
  + `2` or `"all"`: 1/2 x `loq` 

Setting `loq_method = 1` is useful for plots on a linear scale, 
where pre-dose concentrations can be zero. A horizontal reference 
line is drawn at `y=loq` with a corresponding legend entry and caption.
                        
```{r plot-dvtime-blq1, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_dvtime(plot_data, col_var = Dose,
            loq_method = 1, loq = 1) +
  scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") 
```

{{< pagebreak >}}

Setting `loq_method = 2` is useful for visualizing plots on a log scale, 
where pre-dose concentrations are best visualized imputed to 1/2 LLOQ. The 
horizontal reference line and legend entries remain; however, the caption
updates to define the new imputation.

```{r plot-dvtime-blq2-log, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_dvtime(plot_data, col_var = Dose, log_y = TRUE,
            loq_method = 2, loq = 1) +
  scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") 
```

{{< pagebreak >}}

Dose-normalization in `plot_dvtime` is also handled in a standardized manner, 
without modifying the underlying data for visualization. Like BLQ imputation, 
dose-normalization requires specification of two arguments: `dose_var` and
`dosenorm`.

The `dose_var` argument specifies the variable containing the numeric value of 
dose on observation records (default is `DOSE`). This argument supports 
non-standard evaluation, and may be specified as a bare name or as a string.

The `dosenorm` argument is a logical specifying if dose normalization should be
applied.

```{r plot-dvtime-dosenorm, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_dvtime(plot_data, col_var = Dose, log_y = TRUE,
            dosenorm = TRUE) +
  scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "Dose-normalized Concentration (ng/mL)", x = "Time (hours)") 
```