---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
options(scipen = 999)
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(patchwork)
library(mrgsolve)
library(pmxhelpr)
```

```{r data-sad-pk, eval=T, echo=F, include=F}
data_sad_pk <- data_sad %>% 
  filter(CMT %in% c(1,2)) %>% 
  filter(FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

```{r model-pk, eval=T, echo=F, include=F}
model <- model_mread_load("pkmodel")
```

```{r df_mrgsim_replicate, eval=T, echo=F, include=F}
simout <- df_mrgsim_replicate(data = data_sad_pk,
                     model = model,
                     replicates = 500,
                     carry_out = c("DOSE", "FOOD", "BLQ", "LLOQ"),
                     recover  = c("PART", "Dose"))
```

Many VPC tools use algorithmic binning of continuous continuous longitudinal 
data without leveraging the fact that most non-clinical and clinical 
study data include information on exact nominal time of sampling (e.g., study
visit and timepoint). Algorithms ignoring this structure may not adequately 
reproduce the sampling schema in the underlying trial, particularly if BLQ data
are imbalanced and not handled properly.

The VPC plotting toolchain for continuous longitudinal data within `pmxhelpr`
consists of a summary function, `df_vpcstats()`, and two dual-mode plotting 
functions, `plot_vpc_cont()` and `plot_vpc_cens()`. The summary `vpc_stats`
object contains all relevant summary statistics within each bin. Plotting 
functions can perform one-step summary statistic calculation and plotting, or
they can use `vpc_stats` objects in a two-step workflow to generate multiple 
plot types with identical grouping.

The entry point into the plotting pipeline is a `data.frame` of simulated 
replicates of the model development dataset (default format is the output from
`df_mrgsim_replicate()`). This example uses a dataset (`simout`) consisting 
of `N = 500` replicates of a single ascending oral dose study with 
administration in the fasted state. 

A one-step VPC plot by `Dose` a semi-log scale is visualized below.
                        
```{r plot-vpc-cont-dose-log, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_obj_dose <- plot_vpc_cont(data = simout, strat_var = Dose)  
plot_obj_dose + scale_y_log10(guide = "axis_logticks") + scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") 
```

{{< pagebreak >}}

Like `plot_dvtime()`, default BLQ handling methods are applied when the 
variable `LLOQ` is present in `data`, or a single numeric value is specified 
using the `loq` argument.

A core functionality of the `pmxhelpr` VPC workflow is to standardize handling
of BLQ data within the functions to avoid the requirement for pre-processing of 
simulated data. The expectation is that input `data` contains simulated 
values at all sampling times (`EVID=0`and `MDV=0|1`) with censoring of observed 
values denoted by `MDV=1`.

The handling of observed and simulated BLQ data in VPCs is as follows:

  + Censored Observations (`OBSDV`): `MDV=1` or `OBSDV<loq`
    - VPC: Censor values | Rank censored values in quantile stats 
    - pcVPC: Censor values | Drop censored values in quantile stats
    
  + Censored Simulated Values (`SIMDV`): `SIMDV<loq`
    - VPC: No censoring | Rank raw values in quantile stats 
    - pcVPC: Censor values | Drop censored values in quantile stats
    
A one-step VPC plot by `Dose` on linear scale is visualized below.
`plot_vpc_cont()` uses `facet_wrap()` internally to build the facet layers, and 
the default `scales="fixed"` is applied and must be over ridden. This will 
throw a warning, as additional facet layers should *NOT* be added in general, 
unless the faceting aesthetics are identical in both layers (e.g., `Dose`). 

```{r plot-vpc-cont-dose-lin, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_obj_dose +  scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") +
  facet_wrap(vars(Dose), scales = "free_y")
```

{{< pagebreak >}}

One-step pcVPC plots pooled across doses are visualized below on both 
linear and semi-log scales. Both observed and simulated values < `loq` are
censored and dropped in calculation of summary statistics. 

In their seminal paper introducing the pcVPC method (PMID: 21302010), 
Mats Karlsson's group stated the following  regarding the handing of BLQ data
in these diagnostics:

"Observations censored due to limit of detection will therefore need to be 
handled in the same way as previously suggested for drop-out censoring of 
observations (22). This approach is based on omitting censored observations 
(e.g., BQL) in a similar manner for both the observed and simulated datasets 
before calculating the VPC statistics. In this case, the interpretation of the
pcVPC/pvcVPC for the continuous observations will be heavily dependent on an 
accurate prediction of the data censoring. Therefore, diagnostics ensuring 
accurate prediction of censoring frequency (e.g., percentage of observation 
below LLOQ) versus the independent variable will be important before assessing 
the pcVPC/pvcVPC for the continuous observations."

More on that later!

```{r plot-vpc-cont-pcvpc, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_obj <- plot_vpc_cont(data = simout, pcvpc = TRUE)

plot_pcvpc_lin <- plot_obj + scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "PRED-corrected Concentration (ng/mL)", x = "Time (hours)") 

plot_pcvpc_log <- plot_pcvpc_lin + scale_y_log10(guide = "axis_logticks") 

plot_pcvpc_lin + plot_pcvpc_log + 
  plot_annotation(tag_levels = 'A', 
                  title = "pcVPCs on Linear (A) and Semi-log (B) Scales")
```

