---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
options(scipen = 999)
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(patchwork)
library(mrgsolve)
library(pmxhelpr)
```

```{r data-sad-pk, eval=T, echo=F, include=F}
data_sad_pk <- data_sad %>% 
  filter(CMT %in% c(1,2)) %>% 
  filter(FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

```{r model-pk, eval=T, echo=F, include=F}
model <- model_mread_load("pkmodel")
```

```{r df_mrgsim_replicate, eval=T, echo=F, include=F}
simout <- df_mrgsim_replicate(data = data_sad_pk,
                     model = model,
                     replicates = 500,
                     carry_out = c("DOSE", "FOOD", "BLQ", "LLOQ"),
                     recover  = c("PART", "Dose"))
```

The VPC plotting toolchain for continuous longitudinal data within `pmxhelpr`
consists of a summary function, `df_vpcstats()`, and two dual-mode plotting 
functions, `plot_vpc_cont()` and `plot_vpc_cens()`. The summary `vpc_stats`
object contains all relevant summary statistics within each bin. Plotting 
functions can perform one-step summary statistic calculation and plotting, or
they can use `vpc_stats` objects in a two-step workflow to generate multiple 
plot types with identical grouping.

`pmxhelpr` contains a function to pre-process and derive summary statistics 
from the observed and simulated data for VPC plotting: `df_vpcstats()` 

The `df_vpcstats` function handles the first two steps in the VPC plot pipeline:
  
  1. Pre-processing the data to validate arguments, filter to observation 
     records, and handle missing data
  2. Grouping, prediction-correction, and calculation of summary statistics
     
VPC summary statistics are calculated for the input  `data` at the study
replicate-level, and subsequently summarized across replicates to identify 
the non-parametric statistical intervals. 

Outputs are grouped by the variables passed to arguments `ntime_var` and 
`strat_var`. 

Let's generate VPC summary statistics using the dataset `simout` introduced
in the last post, which consists of `N = 500` replicates of a single ascending 
oral dose study with administration in the fasted state. 

```{r df-vpcstats-dose, warning=FALSE}
vpcstats_ntime_dose <- df_vpcstats(
  data = simout,
  strat_var = Dose
)
```
  
A helpful informational message prints to let us know that the `loq` argument 
was automatically inherited from the input `data$LLOQ`.

{{< pagebreak >}}

`df_vpcstats()` returns an S3 `pmx_stats` container with three slots:

  + `stats`: quantile summary statistics of observed and simulated data
  + `obs`:  first replicate observation rows for observation scatter overlay
  + `config`: run configuration (`n_replicates`, `loq`, `strat_var`)
  
The S3 summary method provides a nice compact of the `vpc_stats` object 
dimensions, run configuration, and column groups present in stats. The 
`ntime_var` is renamed to `BIN_MID` internally within the function.

```{r df-vpcstats-dose-summary, warning=FALSE}
summary(vpcstats_ntime_dose)
```

The S3 print method provides the same information with a short head preview of 
the first 3 rows of `x$stats` while the S3 as.data.frame method returns the 
`stats` slot as a plain data.frame.

{{< pagebreak >}}

The `vpc_stats` object always outputs summary statistics for concentration, 
prediction-corrected concentration, and proportion BLQ enabling reuse of the
stats container to generate multiple plots (e.g., VPC, pcVPC, proportion BLQ
VPC).

A standard VPC plot stratified by dose is generated from the `vpc_stats` object
for illustration purposes below.

```{r plot-vpc-cont-dose-log, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_vpc_cont(vpcstats_ntime_dose) + 
  scale_y_log10(guide = "axis_logticks") + 
  scale_x_continuous(breaks = seq(0, 168, 24)) + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") 
```
