---
format:
    html:
       title: '<span style="color:green;">(pmxhelpr)</span> Function Friday'
       embed-resources: true
       self-contained-math: true
    pdf:
       fontsize: 12pt
       geometry:
         - margin=0.75in
       fig-width: 7
       fig-height: 4.2
       include-in-header:
         text: |
           \usepackage{fvextra}
           \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
           \usepackage{xcolor}
           \definecolor{pmxgreen}{HTML}{008000}
---
::: {.content-visible when-format="pdf"}
# `\textcolor{pmxgreen}{(pmxhelpr)}`{=latex} Function Friday
:::

```{r pkgs, eval=T, echo=F, include=F}
options(scipen = 999)
rm(list=ls())
library(devtools)
library(tidyverse)
library(here)
library(patchwork)
library(mrgsolve)
library(pmxhelpr)
```

```{r data-sad-pk, eval=T, echo=F, include=F}
data_sad_pk <- data_sad %>% 
  filter(CMT %in% c(1,2)) %>% 
  filter(FOOD == 0) %>%
  rename(DV = ODV) %>%
  mutate(Dose = var_addn(DOSE, ID, sep = "mg"))
```

```{r model-pk, eval=T, echo=F, include=F}
model <- model_mread_load("pkmodel")
```

```{r df_mrgsim_replicate, eval=T, echo=F, include=F}
simout <- df_mrgsim_replicate(data = data_sad_pk,
                     model = model,
                     replicates = 500,
                     carry_out = c("DOSE", "FOOD", "BLQ", "LLOQ"),
                     recover  = c("PART", "Dose"))
```

The VPC plotting toolchain for continuous longitudinal data within `pmxhelpr`
consists of a summary function, `df_vpcstats()`, and two dual-mode plotting 
functions, `plot_vpc_cont()` and `plot_vpc_cens()`. The summary `vpc_stats`
object contains all relevant summary statistics within each bin. Plotting 
functions can perform one-step summary statistic calculation and plotting, or
they can use `vpc_stats` objects in a two-step workflow to generate multiple 
plot types with identical grouping.

The elements displayed in the VPC can be controlled using the `shown` argument.
The helper function `plot_vpc_shown()` is provided to update individual plot 
elements; the defaults can be viewed by calling this function with no arguments.

```{r df-vpcstats-dose, warning=FALSE}
paste(names(plot_vpc_shown()), plot_vpc_shown(), sep = ":")
```

The components of the list correspond to the following vpc plot elements:

-   Observed points: `obs_point`
-   Observed quantile lines: `obs_pi_line`
-   Observed median line: `obs_median_line`
-   Simulated quantile lines: `sim_pi_line`
-   Simulated quantile CIs: `sim_pi_ci`
-   Simulated quantile PI area: `sim_pi_area`
-   Simulated median line: `sim_median_line`
-   Simulated median CI: `sim_median_ci`

{{< pagebreak >}}

One common alternative VPC to the standard 90% non-parametric CIs of the 5th, 
50th, and 95th percentiles is to generate a 90% non-parametric prediction 
interval (PI) spanning the area between the 5th and 95th percentiles - 
especially with limited data where there may not be enough resolution for 
separate confidence intervals of each quantile.

```{r plot-vpc-dose-pi90, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
shown_pi90 <- plot_vpc_shown(sim_median_ci = FALSE, sim_pi_ci = FALSE, 
                             sim_pi_area = TRUE, sim_median_line = TRUE,
                             obs_pi_line = FALSE)
plot_dose_pi90 <- plot_vpc_cont(simout, pcvpc = FALSE, strat_var = Dose, shown = shown_pi90) + 
  scale_x_continuous(breaks = seq(0, 168, 24), limits = c(0, 168)) + 
  scale_y_log10(guide = "axis_logticks") + 
  labs(y = "Concentration (ng/mL)", x = "Time (hours)") 
plot_dose_pi90
```
{{< pagebreak >}}

With the updates to what is shown in the VPC, it is not immediately intuitive
what each element shown represents. `pmxhelpr` includes an additional VPC
plotting function that generates a legend for the plot, which can be combined
into the plot using `patchwork`. 

Calling this function with no arguments produces a legend matching the default
VPC elements `shown`.

```{r plot-vpc-leg, fig.height=1, fig.width=6, message=FALSE, warning=FALSE}
plot_vpc_legend()
```

Since we defined our new plot elements by assigning the output of 
`plot_vpc_shown()` to an object, we can simply recycle it to make the legend 
matching our updated plot. An complete figure is then output by using 
`patchwork`to combine the VPC plot with the legend in one figure!

```{r plot-vpc-dose-pi90-leg, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
plot_leg_pi90 <- plot_vpc_legend(shown = shown_pi90, lloq = 1)

plot_dose_pi90 / plot_leg_pi90 + plot_layout(heights = c(2, 1))
```


